---
title: |
  Rural-Urban Labour Migration - **Ghana**
  
  _GHANA STATISTICAL SERVICE_ 
  
  _FLOWMINDER.ORG_
  
  *2025 Report*
author: 
  - Sylvester Agyei-Boadi, Selasi Akaho & Preeti
author-meta: "ONS Accelerator Program"
format: 
  html:
    theme: quart
    css: src/styles.css
    page-layout: full
    # toc: true
    title-block-banner: true
    title-block-style: default
description: "Using aggregted CDR data from Telecel Ghana and Flowminder"
execute: 
  echo: false
  message: false
  warning: false
---


```{r}
#| context: setup
source("setup.R")
```


```{r import, warning=FALSE, message=FALSE}
#| cache: true
#| context: data

## import data
# source("modules/imports.R")
pop_est = read.csv('../../pop_release/data/GHA_pop_estimates_2022_2024.csv')
relocations = read.csv('../../pop_release/data/GHA_reloc_estimates_2020_2024.csv') %>%
  mutate(date = as.Date(date))

pop_projection = read.csv('data/Population Projections_20250214-151840.csv')

current_district = 'Achiase'
current_month = as.Date("2024-06-01")
pop_est %>%
  filter(district == current_district) %>%
  slice_head(n =1) %>%
  pull(region) -> current_region

pop_est %>%
  filter(district == current_district) %>%
  slice_head(n = 1) %>%
  pull(dist_code) -> code

geo_data = st_read('data/gss_phc2021_bnd_edgematched_20240821151831.gpkg',
                   layer = "gss_phc2021_16_regions", quiet = TRUE) %>%
  st_make_valid() %>%
  select(!region) %>%
  rename(region = region_name)

```


<!-- :::{.column-screen style="background-color: #f0f8ff; height: 600px; overflow-y: scroll;"} -->
<!-- :::{.column-page} -->
<!-- ::: grid -->
<!-- :::{.g-col-5} -->
:::{.column-screen style="background-color: #F2F2F2; height: 600px; overflow-y: scroll;"}
:::{.column-page}
::: grid 
:::{.g-col-5 style="overflow-y: scroll;"}


```{r defender, output = 'asis'}
defender <- knitr::knit_child("./modules/national/text/introduction.qmd",
                              quiet = TRUE)

cat(defender, sep = "/n")

source("modules/national/charts.R")

```

:::

:::{.g-col-7 style="position: sticky; top: 0; height: 100vh; overflow-x: auto;"}
## Summary Statistics 

<div class="custom-grid">
  <div class="custom-grid-item">
    `r scales::comma(gh_total)`
  Ghana's total dynamic population
  </div>
  <div class="custom-grid-item">
    Monthly change
  </div>
  <div class="custom-grid-item">
    Year-on-year change
  </div>
  <div class="custom-grid-item">
    Difference from projection
  </div>
</div>
 
<br>

```{r gh_trend}
#| fig-height: 3

source("modules/national/charts.R")

gh_trend
```

:::
:::
:::
:::

## Results
### Regional Population and Comparism to Projections

Greater Accra followed by Ashanti had the highest population of 6181206 and 5670095 respectively. Where as Greater Accra followed by Ashanti had the lowest population of 6181206 and 5670095 respectively.

Upper West, Bono, Oti, Northern, Bono East, Western, Ashanti and North East lost population compared to regional projections.

```{r}
#| label: nat
#| column: screen-inset-shaded
#| layout-nrow: 1

regional_map

projected_diff
```


### Relocations Between Regions

In this section we focus on people's relocations, i.e. the estimated number of persons relocating
(i.e. changing their home location) between regions. Each relocation measurement is
initially calculated using the monthly median value of all people relocating, in the period December 2024
and the area of interest. We look first at the largest relocations between regions, and then separately at relocations from and to areas of interest

```{r unsued}

# # make regional map
# reg_data <- pop_est %>%
#   mutate(date = as.Date(date)) %>%
#   filter(date == current_month) %>%
#   group_by(region) %>%
#   summarise(population = sum(pop_est_cdr)) %>% 
#   right_join(geo_data, by = "region") %>% 
#   select(!population) %>% 
#   st_as_sf()
# 
# # Function to create curved lines with gradient colors
# create_curved_gradient <- function(x1, y1, x2, y2, curvature = 0.2, n = 100) {
#   # Calculate control points for the Bezier curve
#   mid_x <- (x1 + x2) / 2
#   mid_y <- (y1 + y2) / 2
#   control_x <- mid_x + curvature * (y2 - y1)
#   control_y <- mid_y - curvature * (x2 - x1)
#   
#   # Generate points along the Bezier curve
#   t <- seq(0, 1, length.out = n)
#   x <- (1 - t)^2 * x1 + 2 * (1 - t) * t * control_x + t^2 * x2
#   y <- (1 - t)^2 * y1 + 2 * (1 - t) * t * control_y + t^2 * y2
#   
#   tibble(
#     x = x,
#     y = y,
#     gradient = t  # Gradient from 0 (origin) to 1 (destination)
#   )
# }

# ## import data
# # source("modules/imports.R")
# pop_est = read.csv('../../pop_release/data/GHA_pop_estimates_2022_2024.csv')
# relocations = read.csv('../../pop_release/data/GHA_reloc_estimates_2020_2024.csv') %>%
#   mutate(date = as.Date(date))
# 
# current_month = as.Date("2024-12-01")
# 
# geo_data = st_read('data/gss_phc2021_bnd_edgematched_20240821151831.gpkg',
#                    layer = "gss_phc2021_16_regions", quiet = TRUE) %>%
#   st_make_valid() %>%
#   select(!region) %>%
#   rename(region = region_name)
# 
# # # make regional map
# regional_data <- pop_est %>%
#   mutate(date = as.Date(date)) %>%
#   filter(date == current_month) %>%
#   group_by(region) %>%
#   summarise(population = sum(pop_est_cdr)) %>%
#   right_join(geo_data, by = "region") %>%
#   select(!population) %>%
#   st_as_sf()
```

```{r}
#| column: page

library(bslib)
library(shiny)

page_fillable(
  layout_columns( 
    card( 
      card_header("Relocations to the region"),
      p("Select a region to display relocations from other regions into the selected region."),
      selectInput("region", "Select Region:",
              choices = unique(relocations$to_region),
              selected = "Greater Accra"),
      tags$img(src = "src/img/flow_svg1.svg", alt = "Relocations from other regions into the selected region")
      ),
    card( 
      card_header("Relocations from the region"),
      p("Select a region to display relocations from the selected regions to other region."),
      selectInput("region", "Select Region:",
              choices = unique(relocations$to_region),
              selected = "Greater Accra"),
      tags$img(src = "src/img/outflow_svg.svg", alt = "Relocations from the selected region to other regions")
      ) 
  ) 
)

```


<!-- ::: {.columns} -->

<!-- ::: {.column width="100%"  style="text-align: center;"} -->
<!-- **Relocations from the region** -->
<!-- ```{r} -->
<!-- #| label: reg -->

<!-- remove(regional_map, pop_projected, pop_palette, projected_diff, reg_palette) -->

<!-- source("modules/regional/reg_charts.R") -->

<!-- print(inflow_to_reg) -->
<!-- ``` -->
<!-- ::: -->


<!-- ::: {.column width="50%"  style="text-align: center;"} -->
<!-- **Relocations to the region** -->
<!-- ```{r} -->
<!-- #| label: reg2 -->

<!-- print(inflow_to_reg) -->
<!-- ``` -->
<!-- :::  -->

<!-- ::: -->





